---
title: "seminribm package: SEmInR model, stochastic, individual-based."
author: "David Champredon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{seminribm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `seminribm` package implements an epidemiological model of type SEIR (susceptible, exposed, infectious and recovered). The model is individual based and stochastic. The stochastic algorithm follows a standard Gillepsie implementation, with or without a tau-leap approximation.


# Example
Let's start by defining the model parameters:
```{r param_setup, echo=TRUE,message=FALSE, warning=FALSE}

horizon <- 300
popSize <- 5e3

initInfectious  <- 2
R0              <- 3.0
latent_mean     <- 2
infectious_mean <- 4
nE              <- 6
nI              <- 6
calc_WIW_Re     <- FALSE
doExact         <- FALSE
timeStepTauLeap <- 0.1
rnd_seed        <- 1234
```

We can now run the model:

```{r run_simulation}
library(seminribm)
res <- seminribm_run(horizon,
                     popSize ,
                     R0 ,
                     latent_mean ,
                     infectious_mean,
                     nE ,
                     nI ,
                     initInfectious ,
                     doExact ,
                     timeStepTauLeap,
                     rnd_seed ,
                     calc_WIW_Re)
```

The time series of the number of susceptible, infectious and recovered individuals can be plotted:
```{r plot_timeseries, fig.width=7, fig.height=7}
plot(res$times, res$S, 
     typ='l', las=1,
     ylim=range(res$S,res$prev),
     main = 'Epidemic Time Series',
     xlab = 'time', ylab = 'n')
grid()
lines(res$times, res$prev, typ='l', col='red2')
lines(res$times, res$R, typ='l', col='green3')
```

We can also extract information about generation intervals. The following figure plots the backward generation intervals as a function of infectee's disease acquisition time:

```{r gi_bck, fig.width=7, fig.height=7, message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2) ; theme_set(theme_bw())

# Retrieve backward GIs and acquiitoin times:
b <- res$GI_bck
at <- res$acq_times
    
# Round times to average per time unit:
df.b <- data.frame(at,b) %>%
    mutate(rt = round(at))

# Plot:
df.b %>%
    group_by(rt) %>%
    summarise(b_mean = mean(b)) %>%
    ggplot(aes(x=rt,y=b_mean)) + 
    geom_point(data = df.b, 
               aes(x=at,y=b), 
               alpha=0.15, colour="orange",
               pch=16,size=4) +
    geom_abline(slope = 1, intercept = 0, linetype=2, colour = "grey")+
    geom_line(size=1.5) + 
    geom_point(size=2) +
    ggtitle('Backward Generation Intervals (line: daily mean)')+
    xlab('calendar time')+ylab('days')

```

Similarly, the _forward_ generation intervals can be plotted. 

```{r gi_fwd, fig.width=7, fig.height=7, message=FALSE}
# Unpacking forward GIs 
# and link to acquisition date:
f <- res$GI_fwd
tat <- res$acqTransm_t

tat2 <- vector()
gifwd <- vector()
k <- 1
for(i in 1:length(f)){
    for(j in 1:length(f[[i]])){
        tat2[k] <- tat[i]
        gifwd[k] <- f[[i]][j]
        k <- k+1
    }
}

# Round time for later averaging:
df.f <- data.frame(at=tat2, rt=round(tat2), f=gifwd)

df.f2 <- df.f %>%
    group_by(rt) %>%
    summarise(f_mean = mean(f))

ggplot(df.f) + 
    geom_point(aes(x=rt,y=f), alpha=0.1, size=4, pch=16, colour="blue") +
    geom_line(data=df.f2, aes(x=rt, y=f_mean), size=2)+
    ggtitle('Forward Generation Intervals (line: daily mean)')+
    xlab('calendar time')+ylab('days')
```

Finally, we can look at the temporal evolution and distribution of the number of secondary cases:

```{r secondary_cases, fig.width=7, fig.height=7, message=FALSE}
df.R <- data.frame(ta = res$Reff[[1]], 
                   tar = round(res$Reff[[1]]),
                   sc = res$Reff[[2]])

df.R %>% 
    group_by(tar) %>%
    summarize(sc_mean = mean(sc)) %>%
    ggplot(aes(x=tar,y=sc_mean)) + 
    geom_hline(yintercept = R0, linetype=2) +
    geom_point(data=df.R, aes(x=ta, y=sc), 
               alpha=0.2, pch=16,size=3, colour="red")+
    geom_line(size=2) +
    ggtitle('Secondary cases')+
    xlab('time infector disease acquisition')+ylab('num. secondary transmissions')

ggplot(df.R) + 
    geom_histogram(aes(x=sc), binwidth = 1) +
    scale_x_continuous(breaks=0:(max(df.R$sc+1))) +
    ggtitle('Secondary cases')+
    xlab('num. secondary transmissions')

```




